---
title: "R Notebook"
output: html_notebook
---
Created by: Sri Naga Sai Sushma Kondabolu
CS5200 Spring 2022

In the following code chunk connected R to the 'birdstrikes' db created locally on MySQL
```{r}

library(RMySQL)     


db_user <- 'root'
db_password <- '100101'
db_name <- 'birdstrikes' 


db_host <- '127.0.0.1' 
db_port <- 3306 


mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,dbname = db_name, host = db_host, port = db_port)
```

Connecting to database in MySQL using password

```{r}
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,dbname = db_name, host = db_host, port = db_port)
```

Dropping tables if they are existing already
```{sql connection=mydb}
DROP TABLE IF EXISTS general;

```

```{sql connection=mydb}
DROP TABLE IF EXISTS incidents;

```

```{sql connection=mydb}
DROP TABLE IF EXISTS conditions;

```

```{sql connection=mydb}
DROP TABLE IF EXISTS airports;

```


Creating 'airports' table
```{sql connection=mydb}
CREATE TABLE airports(
  aid INTEGER PRIMARY KEY AUTO_INCREMENT,
  airportName VARCHAR(45) UNIQUE,
  airportCode INT DEFAULT NULL,
  state VARCHAR(45)
)

```




Creating 'conditions' table
```{sql connection=mydb}
CREATE TABLE conditions(
  cid INTEGER PRIMARY KEY AUTO_INCREMENT,
  condition_name VARCHAR(45) UNIQUE,
  explanation VARCHAR(45) DEFAULT NULL
)

```



Creating 'incidents' table
```{sql connection=mydb}
CREATE TABLE incidents(
  iid INT PRIMARY KEY AUTO_INCREMENT,
  _date DATE,
  origin INTEGER,
  airline VARCHAR(45),
  aircraft VARCHAR(45),
  flightPhase VARCHAR(45),
  impact BOOLEAN,
  cond INTEGER,
  FOREIGN KEY (origin) REFERENCES airports(aid),
  FOREIGN KEY (cond) REFERENCES conditions(cid),
  CHECK (flightphase = "takeoff" OR flightphase = "landing" 
    OR flightphase = "inflight" OR flightphase = "unknown" )
);

```



Importing data in csv file to a dataframe
```{r}

file_name = "BirdStrikesData.csv"
df = read.csv(file=file_name,check.names=FALSE) 

df$FlightDate = sub("0:00","",df$FlightDate)
df$FlightDate = trimws(df$FlightDate)
df$FlightDate = as.Date(df$FlightDate, format = "%m/%d/%Y")

names(df) <- sub(" ", " ", names(df))
names(df) <- sub("﻿", " ", names(df)) 
names(df) <- trimws(names(df))


df[df == ""] = "unknown" 

df
```

Checking if the tables present in the database created
```{r}
dbListTables(mydb)
```


```{sql connection=mydb}
SET GLOBAL local_infile = true;
```


Writing the dataframe into a table named 'general'
```{r}

dbWriteTable(conn=mydb, name="general", value=df)
```

checking the names of all columns in the table 'general'
```{sql connection=mydb}
SHOW COLUMNS FROM general;
```

Displaying the data in 'general' table
```{sql connection=mydb}
SELECT *  FROM general;

```


Inserting different airports name and their origin state into 'airports' table
```{sql connection=mydb}
INSERT INTO airports(airportName, state)
SELECT DISTINCT `Airport: Name`, `Origin State` 
FROM general
WHERE `Airport: Name` <> "unknown"
```


Adding unknown airport name and state to table
```{sql connection=mydb}
INSERT INTO airports(airportName, state)
VALUES("unknown", "unknown") 

```


Displaying the data inserted into airports table
```{sql connection=mydb}
SELECT * FROM airports

```




Inserting different sky conditions into 'conditions' table

```{sql connection=mydb}
INSERT INTO conditions(condition_name)
SELECT DISTINCT `Conditions: Sky` 
FROM general
```



Inserting data into 'incidents' table, date is displayed in desired format, aid is linked to origin, impact is stored as boolean, flightphase is normalized.

```{sql connection=mydb}
INSERT INTO incidents ( iid, _date, origin, airline, aircraft, impact, flightphase, cond)
SELECT t.`Record ID`, STR_TO_DATE(t.`FlightDate`,'%Y-%m-%d'),
case
when t.`Aircraft: Airline/Operator` is not null then (select a.aid from airports a where t.`Airport: Name`=a.`airportName` and t.`Origin State`=a.`state`)
else (select a.aid from airports a where a.`airportName`='unknown')
end
, t.`Aircraft: Airline/Operator`, t.`Aircraft: Make/Model`,
case 
when t.`Effect: Indicated Damage` like '%Caused damage%' THEN TRUE
ELSE FALSE
end
,
CASE
WHEN t.`When: Phase of flight` LIKE '%Take%' THEN 'takeoff'
WHEN t.`When: Phase of flight` LIKE '%Climb%' THEN 'inflight'
WHEN t.`When: Phase of flight` LIKE "%descent%" THEN 'inflight'
WHEN t.`When: Phase of flight` LIKE '%landing%' THEN 'landing'
WHEN t.`When: Phase of flight` LIKE '%approach%' THEN 'landing'
ELSE 'unknown'
END,
c.cid
FROM general t, conditions c where t.`Conditions: Sky`=c.`condition_name`

```


Displaying data read into incidents
```{sql connection=mydb}
SELECT * FROM incidents

```
```{sql connection=mydb}

delete from incidents where airline='MILITARY';

```

Displaying data loaded from all tables
```{sql connection=mydb}
SELECT count(*) FROM incidents

```
```{sql connection=mydb}
select * from conditions
```




```{sql connection=mydb}
SELECT * FROM airports;

```
```{sql connection=mydb}
SELECT *
FROM incidents
```
Describing the flightPhase, aircraft, and airline for each strike, to see how different airlines or aircrafts experienced strike in various flight phases.
```{sql connection=mydb}

select iid, airline, aircraft, flightPhase from incidents ;

```
Viewing all condition names in conditions table to check the different type of conditions.
```{sql connection=mydb}

select condition_name from conditions

```
Checking names for different states from airports table

```{sql connection=mydb}

select distinct state from airports

```




Count of incidents by each flight phase

```{sql connection=mydb}
select flightPhase, count(iid) from incidents group by flightPhase;
```

Flightphase with above average number of bird strikes
```{sql connection=mydb}
select flightPhase from incidents having count(iid)>(select avg(count) from (select count(*) as 'count' from incidents group by flightPhase)sub)
```
Average bird strikes per month across all years

```{sql connection=mydb}
select month,avg(count) from (select year(_date),month(_date) as 'month',count(*) as 'count' from incidents group by year(_date),month(_date))sub group by month;
```

without null data
```{sql connection=mydb}
select month,avg(count) from (select year(_date),month(_date) as 'month',count(*) as 'count' from incidents group by year(_date),month(_date))sub where month is not null group by month;
```



reference
```{sql connection=mydb}
Select flightPhase,Year_considered, AVG(Incidents_count) FROM (SELECT flightPhase,month(_date) as month_considered,year(_date)as Year_considered,COUNT(*) Incidents_count from incidents
GROUP BY flightPhase,month(_date)) as temp
GROUP BY flightPhase,Year_considered
ORDER BY Year_considered,flightPhase

```


```{sql connection=mydb}

SELECT 
  year(_date) as "Year",
  COUNT(year(_date)) as "Strike Incident Count"
FROM incidents where year(_date)>2004 and year(_date)<2012
group by year(_date) order by year(_date)

```
```{r}

library(sqldf)
value <-dbGetQuery(mydb,"SELECT 
  year(_date) ,
  COUNT(year(_date)) 
FROM incidents where year(_date)>2004 and year(_date)<2012
group by year(_date) order by year(_date)")



names(value)[names(value) == "year(_date)"] <- "Year"

names(value)[names(value) == "COUNT(year(_date))"] <- "Count"
value
```
Plotting of column chart with the help of data frame created above
```{r}
library(ggplot2)
ggplot(data=value, aes(x=Year, y=Count)) +
    geom_bar(stat="identity")+ggtitle('Count of bird strikes vs Year')
```


Procedure to add birdstrike

```{sql connection=mydb}
DROP PROCEDURE IF EXISTS add_bird_strike


```


insert into airports(airportName,state) values (origin_airport,origin_state) where origin_airport not in (select airportName from airports) and origin_airport<>'';
```{sql connection=mydb}

CREATE PROCEDURE add_bird_strike(origin_airport VARCHAR(50), origin_state VARCHAR(50),flight_phase VARCHAR(50),impact_bool VARCHAR(50),condition_sky VARCHAR(50))

begin

insert into airports(airportName,state)
select origin_airport,origin_state where origin_airport not in (select airportName from airports) and origin_airport<>'';



INSERT INTO incidents(origin,impact, flightPhase,cond)
values (
case
when origin_airport in (select airportName from airports) then (select a.aid from airports a where a.airportName=origin_airport)
else (select a.aid from airports a where a.`airportName`='unknown')
end,
case 
when impact_bool like '%Caused damage%' THEN TRUE
ELSE FALSE
end,
CASE
WHEN flight_phase LIKE '%Take%' THEN 'takeoff'
WHEN flight_phase LIKE '%Climb%' THEN 'inflight'
WHEN flight_phase LIKE "%descent%" THEN 'inflight'
WHEN flight_phase LIKE '%landing%' THEN 'landing'
WHEN flight_phase LIKE '%approach%' THEN 'landing'
ELSE 'unknown'
END
,
(select cid from conditions where condition_name=condition_sky)




);


END

```
```{sql connection=mydb}
call add_bird_strike('Sushma','Florida','takeoff','Caused damage','No Cloud')
```


Inserted above record
```{sql connection=mydb}

select * from incidents,airports where incidents.origin=airports.aid and airports.airportName='Sushma'
```



